MLH1 Results Outline

### 2020-02-21 doc/Results.rmd
organizing the Q2 IFD section

total SC t test

 p = `r Q2tot.sc_t.test_High.Low$p.value`
**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

musc and mol ttests

p = `r`

p= `r`

@@

@@

@@

@@

In order to test /

@@

<!-- set up chunks here

```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
#isolate all musc
#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

@@

```{r Q2.show.nrmF1, echo=FALSE}
#Curated_BivData_male

@@

Remade the Position table. and MSM has a more telomeric distribution. MOLF has the most central distribution.

```{r Q2.F1.pos.LONG.BIV, echo=FALSE}

@@

A main / biggish cavest to address for this section is the chromosome size effect, use real.long.bivData.

or have chromosome lenght as a effect? would only work for models using the single bivalents, not mouse averages.

<!-- 
Main DFs for this section:
Male.poly.Mouse.Table_BivData_4MM
male.bivdata.1CO
-->

@@

The t.tests are not significant for normalized foci position comparing all bivalets

The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).

@@

mouse mean
short bivalent

So try using the bivalent sample levels

@@

The t.test at the bivalent level is also NS (test; p = )

@@

@@

The normalized 1CO position is not a good metric to distinguish the high and low rec groups.

GLMs

@@

```{r Q2.1CO.landscape, echo=FALSE}
#Male.poly.Mouse.Table_BivData_4MM
#Curated_BivData_male

@@

## JUNK Section

@@

@@

B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.

@@

@@

@@

@@

```{r Q2.IFD.PER, echo=FALSE}

### Q2 Long Bivalent Dataset

```{r Q2.long_biv.IFD, echo=FALSE}
#long BivData
male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)

table(droplevels(male.long.biv.2CO_IFD$strain) )

The plot above shows the normalized IFD measures from 2CO bivalents

There are many fewer long bivalent data observations, but the same positive correlation with

-table of long bivalents

- glms

- logistic regression

@@

There are not enough observations for sub setting the data into strains or use the mouse averages.

Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
The above t.tests are breaking the knitr

None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.

@@

#Caveats

put all of the code chunks/analysis for caveats here

## Chromosome Size Effect

I tried to isolate bivalents which are in the top quartille for SC length from their cells. (re think where this section should go)

<!-- 
Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
--> 
Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plots excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correctly mapped)

@@

                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
        #add Q4

@@

```{r long.biv.plots, echo=FALSE}

@@

## Adjusting for XX

@@

#### Whole Cell Manual Comparison

>>>>

total SC t test
 p = `r Q2tot.sc_t.test_High.Low$p.value`
**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

musc and mol ttests
p = 
p=

@@

<Compare by chrm class, but only for strains with similar proportions>

@@

<lots of variation across SC lengths -- doesn't fit clean expectation or signal>
<balancing the interpertation of general positive correlation of CO number and SC length (SC length drives CO number) --- yet the pattern breaks down -it's not completely predictive -- /at finer levels?>  <noise/mouse effects/chromosome identity effects>

<just like the sex differences in SC area -- this points to differential usage of the SC for placing CO -- investigate the recombination landscape>

@@

# breief transition, rec landscape

the different lack of predictive power of SC length to predict CO number -- indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.

@@

<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents  -->

@@

<!-- set up chunks here -->

```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
#isolate all musc
#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

@@

```{r Q2.1CO.show_plots, echo=FALSE}
#Curated_BivData_male

@@

```{r Q2.1COpos.LONG.BIV, echo=FALSE}

@@

<!-- t test -->

@@

@@

@@

@@

The t.tests are not significant for normalized foci position comparing all bivalets
The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).

@@

For normalized position of 1CO. 
The normalized 1CO position is not a good metric to distinguish the high and low rec groups.

<!-- GLMs -->

<toal data set, short and long biv>

@@

```{r Q2.1COpos.glm, echo=FALSE}
#Male.poly.Mouse.Table_BivData_4MM
#Curated_BivData_male

@@

<no prediction for the strains within subspecies>

@@

```{r Q2.IFD.long.short.setup, echo=FALSE}
#long BivData
male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)

@@

@@

<motivation background / fit more COs along a chromsome>

t.test, are the high and low rec strains different?

-glms -- are predictions based on the gwRR met for interference

@@

Full_2CO dataset
- short biv data
- long bivdata

<glms>

@@

```{r Q2.IFD.PER.glm, echo=FALSE}

@@

@@

<None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.>

@@

<!-- 
Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
-->

<Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plots excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correctly mapped)>

@@

                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)

@@

<below chunk goes with the above chunk>
```{r long.biv.plots, echo=FALSE}

@@

# Adjusting for XX

@@

# Whole Cell Manual Comparison
